### 2015-10-15 manuscript.Rnw
small edits to the text

Phylogenetic information retrieved from OTL is converted into
\texttt{ape::phylo} objects by \texttt{rotl} using the NEXUS Class Library
\citep[NCL,][]{Lewis2003} as implemented in the \texttt{rncl} package
\url{https://cran.r-project.org/package=rncl}. Using NCL provides robust and
efficient parsing of large trees that may contain singleton nodes labeled with
taxonomic information.

@@

The function \texttt{tnrs\_match\_names} returns a data frame that lists the
Open Tree identifiers as well as other information to help users to ensure that
the taxa matched are the correct ones. Here, there is no ambiguity in the taxa
matched, however, as the Open Tree Taxonomy includes taxa from bacteria, plants,
and animals that are regulated by different nomenclatural codes (ICNP, ICN and
ICZN respectively), OTL and \texttt{rotl} provide tools to deal with potential
hemihomonyms. The argument \texttt{context\_name} can be used to limit potential
matches to a taxonomic group (see the function \texttt{tnrs\_contexts} for a list
of possible options).  When this strategy cannot be used in case the tree
encompasses multiple domains (as in the present example), the function
\texttt{inspect} lists alternative matches for a taxon name, and \texttt{update}
replaces it in the results. An example of this approach is provided in the
vignette ``How to use \texttt{rotl}?'' that accompanies the package.

@@

Criteria that can be used to search for studies or their associated trees are
available through the output of the function
\texttt{studies\_properties}. Typically, users will want to search for studies
or trees based on taxon names (or their Open Taxonomy identifiers), but other
criteria such as the title of the publication can be used. Here we demonstrate
how to look for and retrieve a tree for study focusing on the family Felidae
(Figure~\ref{fig:plot_cats}).

@@

In recent years, R has become part of the toolbox of many researchers in
evolutionary biology. As many methods for comparative analyses are implemented,
and because it is a relatively easy to use programming language, R greatly
facilitates the analysis of large datasets, and allows researchers to combine
methods in novel ways. Additionally, as more data is made available online and
accessible using APIs, many R packages have been developed to take advantage of
these datasets, especially spearheaded by the organization rOpenSci
\url{https://ropensci.org}. \texttt{rotl} contributes to these initiatives and
data from OTL can be combined with other sources easily. For instance in
Figure~\ref{fig:cat_map}, we show how we can obtain a map of the occurrences for
some of the cat species (genus \textit{Lynx}) included in the phylogeny above
using records for these species found in GBIF (code is available in
supplementary material).

>>>>

Phylogenetic information retrieved from OTL is converted into
\texttt{ape::phylo} objects by \texttt{rotl} using the NEXUS Class Library
\citep[NCL,][]{Lewis2003} as implemented in the \texttt{rncl} package
\url{https://cran.r-project.org/package=rncl}. Using NCL provides robust and
efficient parsing of large trees that may contain singleton nodes labeled with
taxonomic information (i.e., monotypic taxon).

@@

The function \texttt{tnrs\_match\_names} returns a data frame that lists the
Open Tree identifiers as well as other information to help users ensure that the
taxa matched are the correct ones. Here, there is no ambiguity in the taxa
matched, however, as the Open Tree Taxonomy includes taxa from bacteria, plants,
and animals that are regulated by different nomenclatural codes (ICNP, ICN and
ICZN respectively), OTL and \texttt{rotl} provide tools to deal with potential
hemihomonyms. The argument \texttt{context\_name} can be used to limit potential
matches to a taxonomic group such as ``Animals'' (see the function
\texttt{tnrs\_contexts} for a list of possible options).  When this strategy
cannot be used in case the tree encompasses multiple domains (as in the present
example), the function \texttt{inspect} lists alternative matches for a taxon
name, and \texttt{update} replaces it in the results. An example of this
approach is provided in the vignette ``How to use \texttt{rotl}?'' that
accompanies the package.

@@

Criteria that can be used to search for studies or their associated trees are
available through the output of the function
\texttt{studies\_properties}. Typically, users will want to search for studies
or trees based on taxon names (or their Open Taxonomy identifiers), but other
criteria such as the title of the publication can be used. Here we demonstrate
how to look for and retrieve a tree for studies focusing on the family Felidae
(Figure~\ref{fig:plot_cats}).

@@

In recent years, R has become part of the toolbox of many researchers in
evolutionary biology. As many methods for comparative analyses are implemented,
and because it is a relatively easy to use programming language, R greatly
facilitates the analysis of large datasets, and allows researchers to combine
methods in novel ways. Additionally, as more data is made available online and
accessible using APIs, several R packages have been developed to take advantage of
these datasets, especially spearheaded by the organization rOpenSci
\url{https://ropensci.org}. \texttt{rotl} contributes to these initiatives and
data from OTL can be combined with other sources easily. For instance in
Figure~\ref{fig:cat_map}, we show how we can obtain a map of the occurrences for
some of the cat species (genus \textit{Lynx}) included in the phylogeny above
using records for these species found in GBIF (code is available in
supplementary material).
